---
title: "Statistical Inference Course Project"
author: "Jakub Wiatrak"
date: "20 April 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Overview

This is course project report, created as a part of Statistical Inference course on Coursera. The aim of this report is to investigate exponential distribution and compare it with the Central Limit Theorem in terms of mean and variability.


## Simulations

The following code will simulate a 1000 samples of 40 random exponentials, and take the means of those samples. The rate (lambda) would be 0.2.

```{r}
lambda <- 0.2
n <- 40
sims <- 1000

exp_means <- NULL
for (i in 1:sims) exp_means <- c(exp_means, mean(rexp(n, lambda)))
#converting the simulations to data.frame for easier use with ggplot
exp_means <- data.frame(exp_means)
```


### Theoretical mean and empirical mean

Theoretical mean of exponential distribution is equal to 1/lambda. Let's compare that to the mean of our 1000 simulations:
```{r}
#Theoretical mean
1/lambda

#Empirical mean
mean(exp_means[,1])

```

Those means are very close.

### Theoretical variance and empirical variance

The theoretical variance of exponential distribution is (1/lambda)^2. Therefore, according to CLT, the theoretical variance of n random exponentials is (1/lambda)^2/n.
Let's compare the theoretical variance to the empirical one:

```{r}
#Theoretical variance
(1/lambda)^2/n

#Empirical variance
var(exp_means[,1])
```

As we can see, those variances are very close as well.


### Comparison with normal distribution

First, let's take a look the exponential distribution. The following code simulates 1000 random exponentials with rate (lambda) equal to 0.2.

```{r}
library(ggplot2)
exp_dist <- rexp(1000, lambda)
exp_dist <- data.frame(exp_dist)


ggplot(exp_dist, aes(x=exp_dist)) +
        geom_histogram(aes(y=..density..), binwidth = 2) +
        stat_function(fun=dnorm,
                      color="red",
                      size = 2,
                      args=list(mean=1/lambda, sd=1/lambda))

```
As we can see, the distribution is clearly different from normal (indicated by the red curve).

Now let's plot the simulated means of 1000 samples of 40 random exponentials.

```{r}
ggplot(exp_means, aes(x=exp_means)) +
        geom_histogram(aes(y=..density..), binwidth = 0.2) +
        geom_vline(xintercept=mean(exp_means[,1]), size=1, color = "green") +
        geom_vline(xintercept=1/0.2, size=1, color = "red") +
        stat_function(fun=dnorm,
                      color="red",
                      size = 2,
                      args=list(mean=1/0.2, sd=5/sqrt(40)))
```

It is clearly visible, that the distribution is very similar to normal. For comparison, normal distribution has been ploted on the histogram with a red curve.
The red and green vertical lines, indicating theoretical and empirical means, also show how close the distribution is to the theoretical one, indicated by CLT.

## Conclusion

The analysis exemplifies the Central Limit Theorem in practice. It shows that with larger samples the distribution of variables becomes similar to the normal distribution.